A tool set for random number generation on GPUs in R(?)

### 2021-11-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
random exponential finished
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2947 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

TO BE PUT IN APPENDIX OR DELETE? \\
In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}. The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
<<histAndQQ_prepare,eval=FALSE,cache=TRUE>>=
streams <- createStreamsGpu(n =512*128)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="float")
avector<-as.vector(as.matrix(normal_matrix))
@
\begin{figure}[H]
\centering
 <<histAndQQ, echo=TRUE, eval=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
@
\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
\end{figure}

@@

We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU. The \proglang{R} outputs for testing on computer 2 is shown in the following. 
<<deviceinfo, include=FALSE,cache=FALSE>>=
library(kableExtra)
a <- data.frame(
  var1 = c(" ", "Computer1", "Computer2"),
  var2 = c('Host', 'Intel Xenon W-2145 3.7Ghz', 	'VCPU Intel Xenon Skylake 2.5Ghz'),
  var3 = c('Device', 'AMD Radeon VII', 'VGPU Nvidia Tesla V100')
)

knitr::kable(a, col.names = NULL, align='c', caption = "Device information for run-time comparison.\\label{tab:device}") %>%
  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position")
@
% \begin{table}[H]
% \begin{tabular}{ |c|c|c|c|c|c| } 
% \hline
%            & host   &                     device \\ \hline
%  Computer1 & INtel Xenon W-2145 3.7 Ghz & Amd Radeon VII  \\ \hline
%  Computer2 & VCPU Intel Xenon Skylake 2.5ghz & VGPU Nvidia Tesla V100 \\ \hline
% \end{tabular}
% \caption{ Device information for run-time compare.\label{tab:device}}
% \end{table}

The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}. %The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday. The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.

\subsection{Comparing run-time: Month data example}
<<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
data2<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv") 
data2<-data2[,-1]
month<-as.matrix(data2)
rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
Nglobal = c(256,128)
streams <- createStreamsGpu(n =256*128)
@
<<monthdata,eval=TRUE,include=FALSE, cache=TRUE>>=
library(knitr)
library(kableExtra)
knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@

@@

\subsection{Comparing run-time: Week data example}
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
library(latex2exp)
data<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
#exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
data<-data[,-1]
week<-as.matrix(data)
rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
@
<<weekdata, eval=TRUE, include=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@

@@

knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices. Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII. Computer 2 is equipped with	VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position") %>%
  kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
  add_header_above(c(" " = 1, "Computer 1" = 2, "Computer 2" = 2, " " = 1))
@
% \begin{table}[H]
% \centering
% \begin{tabular}{ |c|c|c|c|c|c| } 
%  \hline
%  data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
%  Month & 1e6 & 0.403804 & 0.403507  & 0.403507     \\ \hline
%  Week &  1e7  & 0.0001251 & 0.0001274 & 0.0001274\\ \hline
% \end{tabular}
% \caption{ Summary of p-values obtained on different devices.\label{tab:summary}}
% \end{table}
% 
% 
% \begin{table}[H]
% \centering
% \begin{tabular}{ |c|c|c|c|c|c| } 
%  \hline
%  data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
%  Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
%  Week &  1e7  & 60.19 & 90.38 & 10.82 & 1.90\\ \hline
% \end{tabular}
% \caption{ Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.\label{tab:compare time}}
% \end{table}

Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example. The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot. 
\begin{figure}[H]
\centering
<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
abline(v = observed_m, col = "blue", lwd = 1.5)

hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
abline(v = observed_w, col = "blue", lwd = 1.5)
@
\caption{ Frequency distribution for test statistics from month and week data tables.\label{fig4}}
\end{figure}

@@

\begin{appendix}

\end{appendix}

>>>>

@@

The arguments of the \fct{gpuRandom::fisher.sim} are described as follows:
\begin{itemize}
\itemsep0em 
  \item \code{x}: a vclMatrix.
  \item \code{N}: number of simulation runs.
  \item \code{streams}: streams.
  \item \code{type}: ``double'' or ``float'' or returned test statistics.
  \item \code{returnStatistics}: if TRUE, return all test statistics. 
  \item \code{Nglobal}: global index space. Default is set as $(64,16)$.
\end{itemize}
We show the advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU. The \proglang{R} outputs for testing on computer 2 is shown in the following. 
<<deviceinfo, include=FALSE,cache=FALSE>>=
library(kableExtra)
a <- data.frame(
  var1 = c(" ", "Computer1", "Computer2"),
  var2 = c('Host', 'Intel Xenon W-2145 3.7Ghz', 	'VCPU Intel Xenon Skylake 2.5Ghz'),
  var3 = c('Device', 'AMD Radeon VII', 'VGPU Nvidia Tesla V100')
)

knitr::kable(a, col.names = NULL, align='c', caption = "Device information for run-time comparison.\\label{tab:device}") %>%
  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position")
@
The 2-way contingency tables %\ref{tab:month} and table \ref{tab:week} 
consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.

\subsection{Comparing run-time: Month data example}
<<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
data2<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv") 
data2<-data2[,-1]
month<-as.matrix(data2)
rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
Nglobal = c(256,128)
streams <- createStreamsGpu(n =256*128)
@

@@

\subsection{Comparing run-time: Week data example}
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
library(latex2exp)
data<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
#exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
data<-data[,-1]
week<-as.matrix(data)
rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
@

@@

knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices. Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII. Computer 2 is equipped with	VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position") %>%
  kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
  add_header_above(c(" " = 1, "Computer 1" = 2, "Computer 2" = 2, " " = 1))
@

\begin{figure}[H]
\centering
<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
abline(v = observed_m, col = "blue", lwd = 1.5)

hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
abline(v = observed_w, col = "blue", lwd = 1.5)
@
\caption{Frequency distribution for test statistics from month and week data tables. Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example. The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot. \label{fig4}}
\end{figure}

@@

\begin{appendix}
TO BE PUT IN APPENDIX OR DELETE? \\
In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}. The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
<<histAndQQ_prepare,eval=FALSE,cache=TRUE>>=
streams <- createStreamsGpu(n =512*128)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="float")
avector<-as.vector(as.matrix(normal_matrix))
@
\begin{figure}[H]
\centering
 <<histAndQQ, echo=TRUE, eval=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
@
\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
\end{figure}

The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday. The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
<<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
library(knitr)
library(kableExtra)
knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@

<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@
\end{appendix}
